---
title: "Penguin ANOVA"
author: "Chase Brewster"
date: "11/6/2018"
output: html_document
---

1. Load packages and read-in penguins.csv data  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: 

Is there a significant difference in mean flipper length for female penguins by species? (Chinstrap, Gentoo, Adelie?)

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}
counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )
         ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")
```

- Exploratory graphs
- Summary Statistics
- Levene's tests for equal variances
- one-way ANOVA
- post-hoc testing

```{r}
hist <- ggplot(penguins2, aes(x = flipper_length), bins = 10) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name)

hist

qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~ common_name, scale = "free")

qq

# Leven's test for equal variances (> 2 groups...if you have two groups, use the F-test for equal variances)

# We'll use the function leveneTest from the 'car' package

peng_lev <- leveneTest(flipper_length ~ common_name, data = penguins2)

peng_lev

# We retain the null hypothesis of equal variances (p = 0.06)

penguins_summary <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

penguins_summary
```
##3. One-way ANOVA

Single Factor: common_name
Levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random variabe: flipper_length (mm)

(Always a *single* factor, with *levels* in it, testing a *random variable*)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins?

```{r}

# H0: Mean flipper lengths across all species are equal
# HA: At least 2 means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

# If we take samples from these three groups, and they actually come from populations with the same means, there is almost a 0% chance we would have drawn these means by random chance

# At least two samples were taken from samples with different means. Which ones are different? All three are different form each other? Or something else?

# Post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are significantly different (F(df) = statistic, p < .001,$\alpha$ = 0.05...)

Mean flipper length for Adelie, Gentoo, and Chinstrap penguins are all significantly different from each other with post-hoc Tukey's HSD (stats.....)

